Simulation of angular resolution of a new electromagnetic sampling calorimeter

### 2023-01-16 paper.tex
updates

The $K_{L} \rightarrow \pi^{0}\nu\bar{\nu}$ decay is identified as a final state existing only a single $\pi^{0}$ reconstructed from detected two photons. A hermetic detector system enclosing the decay region should be prepared to confirm only two photons in the final state. There had been an experimental proposal to detect the $K_{L} \rightarrow \pi^{0}\nu\bar{\nu}$ decay with the reconstruction of a single $\pi^{0}$ using directional information of two photons~\cite{KOPIO}. A Preradiator (PR) is placed in front of the calorimeter to measure the photon's incident angle. The PR consists of a close-packed multi-layer sandwich of 8-mm-thick scintillating plastics alternating with the single-plane and this-gap drift chambers with cathode strips. For good energy resolution and high detection efficiency against a few hundred MeV photons, the scintillating plastic is used for electron-positron conversion. This idea had been confirmed by a beam test using tagged photon beams and achieved its angular resolution of 25~mrad for 250~MeV~\cite{KOPIOcdr}. Unfortunately, the proposal was not accepted, and the detector development could not continue.

@@

In this respect, we report on results of Monte Carlo studies on directional measurement analyzing energy-weighted shower shapes generated in a toy sampling calorimeter, which consists of segmented alternating lead and scintillating plastic layers. There is a limitation in the angular resolution from the analysis of shower shape produced in a sampling calorimeter due to the stochastic feature of its generation. A deduction of the incident angle of the EM shower using a lead scintillating fiber sandwiched sampling calorimeter is studied~\cite{PbScint}, and the obtained angular resolution is 0.8 degrees for 100-GeV electrons and expected to be larger than 8 degrees for 1-GeV electrons. We take a machine-learning approach to get better angular resolution from the analysis. Recently, machine-learning analysis has become a critical tool for better measurement of particle identification~\cite{mlp}, timing~\cite{mlt}, and energy~\cite{mle}. We use the XGBoost (XGB), which is an optimized distributed gradient-boosting library~\cite{xgboost:2016}.

@@

\begin{figure}[!hbt]
\centering
\includegraphics[width=0.58\textwidth]{Fig5_width_including7.5_12.5mm.pdf}
\caption{ Angular resolutions are deduced in terms of the strip width for 1-GeV photons at $\theta=$~10$^{\circ}$ }
\label{fig:angle_reco_width}
\end{figure}

To inspect the angular resolution for the spatial resolution of measured EM showers,  %%%
Angular resolutions are deduced by varying the strip width from 5 to 35 mm, as shown in Figure~\ref{fig:angle_reco_width}. The narrower the strip width, the larger is the size of the features. Consequently, the angle-reconstruction performance is better with 10-mm-wide strips than 5-mm-wide strips. However, the wider strips hardly provide detailed information on the EM shower. The 15-mm-wide strips yield a good angular resolution of 1.23~$\pm$~0.01$^{\circ}$,
and are selected for the further study considering limited budget for the actual construction of the detector.   %%%

@@

To further study the feasibility for the detector to be a photon tracker, the reconstruction of the incident angle is conducted with the front layers instead of the full layers. %%
Figure~\ref{fig:angle_reco_layer} (a) shows the reconstructed angular distribution for 1-GeV photons by using only the front 32 layers, which correspond to 6.2$X_{0}$. A fraction of the photons fail to be reconstructed because of insufficient information gathering caused by deeply penetrating photons without shower generation, where the fraction is estimated to be 1.9\%. The failed events are represented as a delta function near 0, and such events are removed by requiring the total energy deposit at the front layers should be higher than 10~MeV, which is 1\% of the incident energy. The angular resolution of the front layer is estimated to be 1.30~$\pm$~0.01$^{\circ}$. The evolution of angular resolution degrades with the reduction of the number of layers used as the number of contributors to the reconstruction decreases.

@@

Figure~\ref{fig:angle_reco_dep_gr} (a) shows the angular resolution as a function of incident angle for different incident energies ($E_{\gamma}$). Note that the front 32 layers are used for the reconstruction of the incident angle. The inefficiency from the usage of the front layers is estimated to be 2\% at zero degree of incident angle and decreases as 0.5\% at 40 degrees of incident angle. The angular resolution does not depend on the incident angle for high incident energies. However, the angular resolution changes significantly for low incident energies at small incident angles. Figure~\ref{fig:angle_reco_dep_gr} (b) shows the angular resolution as a function of the incident photon energy at $\theta=$~25$^{\circ}$. The angular resolution is well fitted by the function $p_{0} \oplus p_{1}/\sqrt{E_{\gamma}(\mathrm{GeV})}$, where $p_{0}$ and $p_{1}$ represent the energy-independent and energy-dependent contributions, respectively, added in quadrature.

@@

Figure~\ref{fig:multi-parameter} shows the angular resolution for different numbers of training samples and different detector configurations. The width of curves represents the statistical uncertainty. All configurations show improvement in the angular resolution with the increasing number of training samples, which indicates that the maturity of machine learning is an important factor in angle reconstruction. With an increase in the number of training events, the angular resolution improves most rapidly for the configuration with 5-mm-wide scintillator strips. In other words, the required number of training events should be large enough to provide correct answer from the correlations 
between many features. The spectra are fitted with the empirical function of $\sqrt{I^{2} + (D/N_{\mathrm{sample}})^{k}}$, where $I$ denotes the intrinsic resolution and $D$ and $k$ are describing the decreasing trend. The numbers of training samples reaching the angular resolution of 1.1 times of $I$ for 25, 15, and 5 mm strips are estimated to be 58k, 280k, and 7.6M, respectively, indicating that the number of training samples are insufficient especially for 5-mm-wide strips.

@@

We report on the optimization of the angle reconstruction of a a toy Lead/Plastic EM sampling calorimeter for photons in the range of 0.1 to 2~GeV, using the machine learning technique for the next KOTO experiment. The hyperparameters of the XGBoost have been tuned to obtain the best angular resolution, in which the correlating effects among them need to be considered. In terms of the detector configuration, strips with narrower width provide better angular resolution with detailed information on the EM shower when the training on the data is fully conducted. However, the configuration with a smaller strip width requires the larger number of channels to have the same coverage, and this negatively affects to train the model. training with respect to the quantity of training data. The angular resolution may be optimized by considering the two competing effects of the strip width and quantity of training data. Focusing on the angle measurement with a compact detector, we can obtain an angular resolution of 1.30~$\pm$~0.01$^{\circ}$ for the configuration of 32 layers (6.2$X_{0}$) of 15-mm-wide strips, which is a reduction of 0.1$^{\circ}$ compared to that of the configuration with full layers (20$X_{0}$) with $10^{5}$ training samples. The angular resolution depends on the incident energy of the photon, and the dependence is expressed as $p_{0} \oplus p_{1}/\sqrt{E_{\gamma}}$. Except for small incident angles ($\theta<$~$^{\circ}$~deg), the two parameters do not significantly change up to $40^{\circ}$, and the dependence is estimated to be 0.24 $\oplus$ 1.24$^{\circ}/\sqrt{E_{\gamma}}$.

>>>>

The $K_{L} \rightarrow \pi^{0}\nu\bar{\nu}$ decay is identified as a final state existing only a single $\pi^{0}$ reconstructed from detected two photons. A hermetic detector system enclosing the decay region should be prepared to confirm only two photons in the final state. There had been an experimental proposal to detect the $K_{L} \rightarrow \pi^{0}\nu\bar{\nu}$ decay with the reconstruction of a single $\pi^{0}$ using directional information of two photons~\cite{KOPIO}. A Preradiator (PR) is placed in front of the calorimeter to measure the photon's incident angle. The PR consists of a close-packed multi-layer sandwich of 8-mm-thick scintillating plastics alternating with the single-plane and this-gap drift chambers with cathode strips. For good energy resolution and high detection efficiency against a few hundred MeV photons, the scintillating plastic is used for electron-positron conversion. This idea had been confirmed by a beam test using tagged photon beams and achieved its angular resolution of 25~mrad for 250~MeV~\cite{KOPIOcdr}.

@@

In this respect, we report on results of Monte Carlo studies on directional measurement analyzing energy-weighted shower shapes generated in a toy sampling calorimeter, which consists of segmented alternating lead and scintillating plastic layers. There is a limitation in the angular resolution from the analysis of shower shape produced in a sampling calorimeter due to the stochastic feature of its generation. A deduction of the incident angle of the EM shower using a lead scintillating fiber sandwiched sampling calorimeter is studied~\cite{PbScint}, and the angular resolution is expected to be larger than 8 degrees for 1-GeV electrons. We take a machine-learning approach to get better angular resolution from the analysis. Recently, machine-learning analysis has become a critical tool for better measurement of particle identification~\cite{mlp}, timing~\cite{mlt}, and energy~\cite{mle}. We use the XGBoost (XGB), which is an optimized distributed gradient-boosting library~\cite{xgboost:2016}.

@@

\begin{figure}[!hbt]
\centering
\includegraphics[width=0.58\textwidth]{Fig5_width_including7.5_12.5mm.pdf}
\caption{ Angular resolutions are deduced in terms of the strip width for 1-GeV photons at $\theta=$~10$^{\circ}$ }
\label{fig:angle_reco_width}
\end{figure}
\subsubsection{strip width}
To inspect the angular resolution for the spatial resolution of measured EM showers,  %%%
Angular resolutions are deduced by varying the strip width from 5 to 35 mm, as shown in Figure~\ref{fig:angle_reco_width}. The narrower the strip width, the larger is the size of the features. Consequently, the angle-reconstruction performance is better with 10-mm-wide strips than 5-mm-wide strips. However, the wider strips hardly provide detailed information on the EM shower. The 15-mm-wide strips yield a good angular resolution of 1.23~$\pm$~0.01$^{\circ}$,
and are selected for the further study considering limited budget for the actual construction of the detector.   %%%

@@

\subsubsection{the number of layers}
To further study the feasibility for the detector to be a photon tracker, the reconstruction of the incident angle is conducted with the front layers instead of the full layers. %%
Figure~\ref{fig:angle_reco_layer} (a) shows the reconstructed angular distribution for 1-GeV photons by using only the front 32 layers, which correspond to 6.2$X_{0}$. A fraction of the photons fail to be reconstructed because of insufficient information gathering caused by deeply penetrating photons without shower generation, where the fraction is estimated to be 1.9\%. The failed events are represented as a delta function near 0, and such events are removed by requiring the total energy deposit at the front layers should be higher than 10~MeV, which is 1\% of the incident energy. The angular resolution of the front layer is estimated to be 1.30~$\pm$~0.01$^{\circ}$. The evolution of angular resolution degrades with the reduction of the number of layers used as the number of contributors to the reconstruction decreases.

@@

\subsubsection{incident energy}
Figure~\ref{fig:angle_reco_dep_gr} (a) shows the angular resolution as a function of incident angle for different incident energies ($E_{\gamma}$). Note that the front 32 layers are used for the reconstruction of the incident angle. The inefficiency from the usage of the front layers is estimated to be 2\% at zero degree of incident angle and decreases as 0.5\% at 40 degrees of incident angle. The angular resolution does not depend on the incident angle for high incident energies. However, the angular resolution changes significantly for low incident energies at small incident angles. Figure~\ref{fig:angle_reco_dep_gr} (b) shows the angular resolution as a function of the incident photon energy at $\theta=$~25$^{\circ}$. The angular resolution is well fitted by the function $p_{0} \oplus p_{1}/\sqrt{E_{\gamma}(\mathrm{GeV})}$, where $p_{0}$ and $p_{1}$ represent the energy-independent and energy-dependent contributions, respectively, added in quadrature.

@@

\subsubsection{training sample}
Figure~\ref{fig:multi-parameter} shows the angular resolution for different numbers of training samples and different detector configurations. The width of curves represents the statistical uncertainty. All configurations show improvement in the angular resolution with the increasing number of training samples, which indicates that the maturity of machine learning is an important factor in angle reconstruction. With an increase in the number of training events, the angular resolution improves most rapidly for the configuration with 5-mm-wide scintillator strips. In other words, the required number of training events should be large enough to provide correct answer from the correlations 
between many features. The spectra are fitted with the empirical function of $\sqrt{I^{2} + (D/N_{\mathrm{sample}})^{k}}$, where $I$ denotes the intrinsic resolution and $D$ and $k$ are describing the decreasing trend. The numbers of training samples reaching the angular resolution of 1.1 times of $I$ for 25, 15, and 5 mm strips are estimated to be 58k, 280k, and 7.6M, respectively, indicating that the number of training samples are insufficient especially for 5-mm-wide strips.

@@

We report on the optimization of the angle reconstruction of a a toy Lead/Plastic EM sampling calorimeter for photons in the range of 0.1 to 2~GeV, using the machine learning technique for the background rejection in the next KOTO experiment. The hyperparameters of the XGBoost have been tuned to obtain the best angular resolution, in which the correlating effects among them need to be considered. In terms of the detector configuration, strips with narrower width provide better angular resolution with detailed information on the EM shower when the training on the data is fully conducted. However, the configuration with a smaller strip width requires the larger number of channels to have the same coverage, and this negatively affects to train the model. training with respect to the quantity of training data. The angular resolution may be optimized by considering the two competing effects of the strip width and quantity of training data. Focusing on the angle measurement with a compact detector, we can obtain an angular resolution of 1.30~$\pm$~0.01$^{\circ}$ for the configuration of 32 layers (6.2$X_{0}$) of 15-mm-wide strips, which is a reduction of 0.1$^{\circ}$ compared to that of the configuration with full layers (20$X_{0}$) with $10^{5}$ training samples. The angular resolution depends on the incident energy of the photon, and the dependence is expressed as $p_{0} \oplus p_{1}/\sqrt{E_{\gamma}}$. Except for small incident angles ($\theta<$~$^{\circ}$~deg), the two parameters do not significantly change up to $40^{\circ}$, and the dependence is estimated to be 0.24 $\oplus$ 1.24$^{\circ}/\sqrt{E_{\gamma}}$.
